```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(sctransform)
library(ggplot2)
library(harmony)
library(ArchR)
library(ggsci)
library(SeuratDisk)
library(harmony)
library(Seurat)
library(ggplot2)
library(patchwork)
library(ggpubr)
library(dplyr)
library(sctransform)
library(pheatmap)
library(Matrix)
library(RColorBrewer)
library(scales)
library(data.table)
library(stats)
library("Nebulosa")
library(ggsci)
library(ArchR)
library(biomaRt)
library(Seurat)
library(ggplot2)
library(patchwork)
library(ggpubr)
library(dplyr)
library(Seurat)
library(patchwork)
library(sctransform)
library(ggplot2)
library(pheatmap)
library(Seurat)
library(ggplot2)
library(Matrix)
library(RColorBrewer)
library(dplyr)
library(scales)
library(data.table)
library(stats)
library("Nebulosa")
```

```{r}
fibroblast <- readRDS("../Fibroblast.dim50.wharmony.filtered.interation3.annotated.rds")
```

```{r}
DimPlot(fibroblast, reduction = 'umap', group.by = 'functional.cluster', label.size = 4, label=FALSE,
        cols = paletteDiscrete(unique(fibroblast$functional.cluster), set = "stallion"))
```

# NHCF
```{r}
NHCF <- readRDS("./NHCF/NHCF.rds")
```

```{r}
anchors_NHCF <- FindTransferAnchors(
  reference = fibroblast,
  query = NHCF,
  normalization.method = "SCT",
  reference.reduction = "spca",
  dims = 1:50
)
```

```{r}
NHCF <- MapQuery(
  anchorset = anchors_NHCF,
  query = NHCF,
  reference = fibroblast,
  refdata = list(
    celltype.l1 = "functional.cluster",
    predicted_CITE = "CITE"
  ),
  reference.reduction = "spca", 
  reduction.model = "umap"
)
```

```{r}
DimPlot(NHCF, reduction = 'ref.umap', group.by = 'predicted.celltype.l1', label.size = 4, label=FALSE,
        cols = paletteDiscrete(unique(fibroblast$functional.cluster), set = "stallion"))

DimPlot(NHCF, reduction = 'umap', group.by = 'predicted.celltype.l1', label.size = 4, label=FALSE,
        cols = paletteDiscrete(unique(fibroblast$functional.cluster), set = "stallion"))
```

```{r}
DimPlot(NHCF, reduction = 'ref.umap', group.by = 'predicted.celltype.l1', label.size = 4, label=FALSE,
        cols = paletteDiscrete(unique(fibroblast$functional.cluster), set = "stallion"), split.by = "group", ncol=3)
```

```{r}
NHCF$predicted.celltype.l1 <- factor(NHCF$predicted.celltype.l1, levels = c("ground state Fib","Myofibroblast","CCL2.THBS1 Fib","APOE Fib","DLK1 Fib","PTGDS Fib","APOD Fib","GDF15 Fib","POSTN+ Fib","Type I IFN Fib","PLA2G2A Fib","POLCE2.MFAP5 Fib","PRG4 Fib"), ordered = TRUE)
```

```{r}
paletteDiscrete(unique(fibroblast$functional.cluster), set = "stallion")
```

```{r}
ggplot(NHCF@meta.data, aes(x=group, fill=predicted.celltype.l1)) + geom_bar(position = "fill") + theme_linedraw() + theme(axis.text.x = element_text(angle = 90)) +  scale_fill_manual(values=c("#D51F26","#272E6A","#208A42","#89288F","#C06CAB","#D8A767","#89C75F")) + theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()) 
```

```{r}
breaksList = seq(0, 1, by = 1/240)
NHCF@assays[["prediction.score.celltype.l1"]]@counts <- NHCF@assays[["prediction.score.celltype.l1"]]@data

DefaultAssay(NHCF) <- "prediction.score.celltype.l1"
Idents(NHCF) <- "predicted.celltype.l1"
sample.averageexpression <- AverageExpression(NHCF,  features = rownames(NHCF@assays[["prediction.score.celltype.l1"]]), assays = "prediction.score.celltype.l1", group.by = "predicted.celltype.l1", slot = "counts")
sample.averageexpression[[1]] <- sample.averageexpression[[1]][c("ground state Fib","Myofibroblast","CCL2.THBS1 Fib","APOE Fib","DLK1 Fib","PTGDS Fib","APOD Fib","GDF15 Fib","POSTN+ Fib","Type I IFN Fib","PLA2G2A Fib","POLCE2.MFAP5 Fib","PRG4 Fib"),]
sample.averageexpression <- as.matrix(sample.averageexpression)[[1]]
pheatmap(sample.averageexpression, scale="none", col=viridis(240), cexCol=0.5, cellwidth=10, cluster_rows=FALSE, fontsize_row=6, fontsize_col=6, cluster_cols = FALSE, legend = TRUE, border_color = NA, cellheight=10,breaks = breaksList)
```

```{r}
FeaturePlot(NHCF, features = c("POSTN+ Fib"),  reduction = "ref.umap", cols = c("lightgrey", "darkred")) & theme(plot.title = element_text(size = 10))
```

```{r}
saveRDS(NHCF, "./NHCF/mapping/NHCF_mapped.rds")
```

# NHDF
```{r}
NHDF <- readRDS("./NHDF/NHDF.rds")
```

```{r}
anchors_NHDF <- FindTransferAnchors(
  reference = fibroblast,
  query = NHDF,
  normalization.method = "SCT",
  reference.reduction = "spca",
  dims = 1:50
)
```

```{r}
NHDF <- MapQuery(
  anchorset = anchors_NHDF,
  query = NHDF,
  reference = fibroblast,
  refdata = list(
    celltype.l1 = "functional.cluster",
    predicted_CITE = "CITE"
  ),
  reference.reduction = "spca", 
  reduction.model = "umap"
)
```

```{r}
DimPlot(NHDF, reduction = 'ref.umap', group.by = 'predicted.celltype.l1', label.size = 4, label=FALSE,
        cols = paletteDiscrete(unique(fibroblast$functional.cluster), set = "stallion"))
```

```{r}
Idents(NHDF) <- "group"
NHDF <- subset(NHDF, idents = c("NHDF_72h_Control","NHDF_72h_IL1b","NHDF_72h_TGFb"))
unique(NHDF$group)
```


```{r}
DimPlot(NHDF, reduction = 'ref.umap', group.by = 'predicted.celltype.l1', label.size = 4, label=FALSE,
        cols = paletteDiscrete(unique(fibroblast$functional.cluster), set = "stallion"), split.by = "group", ncol=3)
```

```{r}
NHDF$predicted.celltype.l1 <- factor(NHDF$predicted.celltype.l1, levels = c("ground state Fib","Myofibroblast","CCL2.THBS1 Fib","APOE Fib","DLK1 Fib","PTGDS Fib","APOD Fib","GDF15 Fib","POSTN+ Fib","Type I IFN Fib","PLA2G2A Fib","POLCE2.MFAP5 Fib","PRG4 Fib"), ordered = TRUE)
```

```{r}
paletteDiscrete(unique(fibroblast$functional.cluster), set = "stallion")
```

```{r}
ggplot(NHDF@meta.data, aes(x=group, fill=predicted.celltype.l1)) + geom_bar(position = "fill") + theme_linedraw() + theme(axis.text.x = element_text(angle = 90)) + scale_fill_manual(values=c("#D51F26","#272E6A","#208A42","#89288F","#FEE500","#8A9FD1","#C06CAB","#D8A767","#90D5E4","#F37B7D")) + theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()) 
```

```{r}
breaksList = seq(0, 1, by = 1/240)
NHDF@assays[["prediction.score.celltype.l1"]]@counts <- NHDF@assays[["prediction.score.celltype.l1"]]@data

DefaultAssay(NHDF) <- "prediction.score.celltype.l1"
Idents(NHDF) <- "predicted.celltype.l1"
sample.averageexpression <- AverageExpression(NHDF,  features = rownames(NHDF@assays[["prediction.score.celltype.l1"]]), assays = "prediction.score.celltype.l1", group.by = "predicted.celltype.l1", slot = "counts")
sample.averageexpression[[1]] <- sample.averageexpression[[1]][c("ground state Fib","Myofibroblast","CCL2.THBS1 Fib","APOE Fib","DLK1 Fib","PTGDS Fib","APOD Fib","GDF15 Fib","POSTN+ Fib","Type I IFN Fib","PLA2G2A Fib","POLCE2.MFAP5 Fib","PRG4 Fib"),]
sample.averageexpression <- as.matrix(sample.averageexpression)[[1]]
pheatmap(sample.averageexpression, scale="none", col=viridis(240), cexCol=0.5, cellwidth=10, cluster_rows=FALSE, fontsize_row=6, fontsize_col=6, cluster_cols = FALSE, legend = TRUE, border_color = NA, cellheight=10, breaks=breaksList)
```

```{r}
FeaturePlot(NHDF, features = c("ground state Fib","Myofibroblast","CCL2.THBS1 Fib","APOE Fib","PTGDS Fib","APOD Fib","GDF15 Fib","POSTN+ Fib","Type I IFN Fib","POLCE2.MFAP5 Fib"),  reduction = "ref.umap", cols = c("lightgrey", "darkred"), ncol = 4) & theme(plot.title = element_text(size = 10))
```

```{r}
saveRDS(NHDF, "./NHDF/mapping/NHDF_mapped.rds")
```

```{r}
NHDF <- readRDS("./NHDF/mapping/NHDF_mapped.rds")
```

```{r}
FeaturePlot(NHDF, features = c("POSTN+ Fib"),  reduction = "ref.umap", cols = c("lightgrey", "darkred")) & theme(plot.title = element_text(size = 10))
```


# iHCF
```{r}
iHCF <- readRDS("./iHCF/iHCF.rds")
```

```{r}
anchors_iHCF <- FindTransferAnchors(
  reference = fibroblast,
  query = iHCF,
  normalization.method = "SCT",
  reference.reduction = "spca",
  dims = 1:50
)
```

```{r}
iHCF <- MapQuery(
  anchorset = anchors_iHCF,
  query = iHCF,
  reference = fibroblast,
  refdata = list(
    celltype.l1 = "functional.cluster",
    predicted_CITE = "CITE"
  ),
  reference.reduction = "spca", 
  reduction.model = "umap"
)
```

```{r}
DimPlot(iHCF, reduction = 'ref.umap', group.by = 'predicted.celltype.l1', label.size = 4, label=FALSE,
        cols = paletteDiscrete(unique(fibroblast$functional.cluster), set = "stallion"))
```

```{r}
DimPlot(iHCF, reduction = 'ref.umap', group.by = 'predicted.celltype.l1', label.size = 4, label=FALSE,
        cols = paletteDiscrete(unique(fibroblast$functional.cluster), set = "stallion"), split.by = "group", ncol=3)
```

```{r}
iHCF$predicted.celltype.l1 <- factor(iHCF$predicted.celltype.l1, levels = c("ground state Fib","Myofibroblast","CCL2.THBS1 Fib","APOE Fib","DLK1 Fib","PTGDS Fib","APOD Fib","GDF15 Fib","POSTN+ Fib","Type I IFN Fib","PLA2G2A Fib","POLCE2.MFAP5 Fib","PRG4 Fib"), ordered = TRUE)
```

```{r}
paletteDiscrete(unique(fibroblast$functional.cluster), set = "stallion")
```

```{r}
ggplot(iHCF@meta.data, aes(x=group, fill=predicted.celltype.l1)) + geom_bar(position = "fill") + theme_linedraw() + theme(axis.text.x = element_text(angle = 90)) + scale_fill_manual(values=c("#D51F26","#272E6A","#208A42","#89288F","#C06CAB","#D8A767","#90D5E4","#89C75F","#9983BD")) + theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.background = element_blank()) 
```

```{r}
breaksList = seq(0, 1, by = 1/240)
iHCF@assays[["prediction.score.celltype.l1"]]@counts <- iHCF@assays[["prediction.score.celltype.l1"]]@data

DefaultAssay(iHCF) <- "prediction.score.celltype.l1"
Idents(iHCF) <- "predicted.celltype.l1"
sample.averageexpression <- AverageExpression(iHCF,  features = rownames(iHCF@assays[["prediction.score.celltype.l1"]]), assays = "prediction.score.celltype.l1", group.by = "predicted.celltype.l1", slot = "counts")
sample.averageexpression[[1]] <- sample.averageexpression[[1]][c("ground state Fib","Myofibroblast","CCL2.THBS1 Fib","APOE Fib","DLK1 Fib","PTGDS Fib","APOD Fib","GDF15 Fib","POSTN+ Fib","Type I IFN Fib","PLA2G2A Fib","POLCE2.MFAP5 Fib","PRG4 Fib"),]
sample.averageexpression <- as.matrix(sample.averageexpression)[[1]]
pheatmap(sample.averageexpression, scale="none", col=viridis(240), cexCol=0.5, cellwidth=10, cluster_rows=FALSE, fontsize_row=6, fontsize_col=6, cluster_cols = FALSE, legend = TRUE, border_color = NA, cellheight=10,breaks = breaksList)
```

```{r}
FeaturePlot(iHCF, features = c("ground state Fib","Myofibroblast","CCL2.THBS1 Fib","APOE Fib","GDF15 Fib","POSTN+ Fib","Type I IFN Fib","PLA2G2A Fib","PRG4 Fib"),  reduction = "ref.umap", cols = c("lightgrey", "darkred"), ncol = 4) & theme(plot.title = element_text(size = 10))
```

```{r}
saveRDS(iHCF, "./iHCF/mapping/iHCF_mapped.rds")
```

```{r}
iHCF <- readRDS("./iHCF/mapping/iHCF_mapped.rds")
```

```{r}
FeaturePlot(iHCF, features = c("POSTN+ Fib"),  reduction = "ref.umap", cols = c("lightgrey", "darkred")) & theme(plot.title = element_text(size = 10))
```


# Fibroblast State Signatures from Brandom/Milena's Paper
# Homeostasis
```{r}
DefaultAssay(NHCF) <- "SCT"
expdata <- GetAssayData(NHCF)
Pop1 <- c("COL1A1","PDGFRA","VIM","DCN")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHCF@meta.data$H1_z<-z_scores[1,]
FeaturePlot(object=NHCF, features = "H1_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

```{r}
DefaultAssay(NHCF) <- "SCT"
expdata <- GetAssayData(NHCF)
Pop1 <- c("WIF1","DKK3","SFRP2","FRZB")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHCF@meta.data$H2_z<-z_scores[1,]
FeaturePlot(object=NHCF, features = "H2_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Injury Response
```{r}
DefaultAssay(NHCF) <- "SCT"
expdata <- GetAssayData(NHCF)
Pop1 <- c("MT1-2","CCL2","CXCL1")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHCF@meta.data$InjuryResponse_z<-z_scores[1,]
FeaturePlot(object=NHCF, features = "InjuryResponse_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Myofibroblast
```{r}
DefaultAssay(NHCF) <- "SCT"
expdata <- GetAssayData(NHCF)
Pop1 <- c("ACTA2","CTHRC1","POSTN","FAP")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHCF@meta.data$Myofibroblast_z<-z_scores[1,]
FeaturePlot(object=NHCF, features = "Myofibroblast_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Matrifibrocyte
```{r}
DefaultAssay(NHCF) <- "SCT"
expdata <- GetAssayData(NHCF)
Pop1 <- c("COMP","CHAD","CLIP2")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHCF@meta.data$Matrifibrocyte_z<-z_scores[1,]
FeaturePlot(object=NHCF, features = "Matrifibrocyte_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Late Resolution
```{r}
DefaultAssay(NHCF) <- "SCT"
expdata <- GetAssayData(NHCF)
Pop1 <- c("COL8A1","MEOX1")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHCF@meta.data$LateResolution_z<-z_scores[1,]
FeaturePlot(object=NHCF, features = "LateResolution_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

```{r}
NHCF@meta.data
```

```{r}
NHCF$group <- factor(NHCF$group, levels = c("NHCF_72h_TGFb", "NHCF_72h_IL1b", "NHCF_72h_Control"))
```

```{r}
VlnPlot(NHCF, features = c("H1_z","H2_z","InjuryResponse_z","Myofibroblast_z","Matrifibrocyte_z"), group.by = "group", pt.size = 0, stack = TRUE, cols = pal_npg(palette = c("nrc"), alpha = 1)(5)) + NoLegend()
```



# Fibroblast State Signatures from Brandom/Milena's Paper
# Homeostasis
```{r}
DefaultAssay(NHDF) <- "SCT"
expdata <- GetAssayData(NHDF)
Pop1 <- c("COL1A1","PDGFRA","VIM","DCN")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHDF@meta.data$H1_z<-z_scores[1,]
FeaturePlot(object=NHDF, features = "H1_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

```{r}
DefaultAssay(NHDF) <- "SCT"
expdata <- GetAssayData(NHDF)
Pop1 <- c("WIF1","DKK3","SFRP2","FRZB")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHDF@meta.data$H2_z<-z_scores[1,]
FeaturePlot(object=NHDF, features = "H2_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Injury Response
```{r}
DefaultAssay(NHDF) <- "SCT"
expdata <- GetAssayData(NHDF)
Pop1 <- c("MT1-2","CCL2","CXCL1")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHDF@meta.data$InjuryResponse_z<-z_scores[1,]
FeaturePlot(object=NHDF, features = "InjuryResponse_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Myofibroblast
```{r}
DefaultAssay(NHDF) <- "SCT"
expdata <- GetAssayData(NHDF)
Pop1 <- c("ACTA2","CTHRC1","POSTN","FAP")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHDF@meta.data$Myofibroblast_z<-z_scores[1,]
FeaturePlot(object=NHDF, features = "Myofibroblast_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Matrifibrocyte
```{r}
DefaultAssay(NHDF) <- "SCT"
expdata <- GetAssayData(NHDF)
Pop1 <- c("COMP","CHAD","CLIP2")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHDF@meta.data$Matrifibrocyte_z<-z_scores[1,]
FeaturePlot(object=NHDF, features = "Matrifibrocyte_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Late Resolution
```{r}
DefaultAssay(NHDF) <- "SCT"
expdata <- GetAssayData(NHDF)
Pop1 <- c("COL8A1","MEOX1")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
NHDF@meta.data$LateResolution_z<-z_scores[1,]
FeaturePlot(object=NHDF, features = "LateResolution_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

```{r}
NHDF$group <- factor(NHDF$group, levels = c("NHDF_72h_TGFb", "NHDF_72h_IL1b", "NHDF_72h_Control"))
```

```{r}
VlnPlot(NHDF, features = c("H1_z","H2_z","InjuryResponse_z","Myofibroblast_z","Matrifibrocyte_z"), group.by = "group", pt.size = 0, stack = TRUE, cols = pal_npg(palette = c("nrc"), alpha = 1)(5)) + NoLegend()
```



# Fibroblast State Signatures from Brandom/Milena's Paper
# Homeostasis
```{r}
DefaultAssay(iHCF) <- "SCT"
expdata <- GetAssayData(iHCF)
Pop1 <- c("COL1A1","PDGFRA","VIM","DCN")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
iHCF@meta.data$H1_z<-z_scores[1,]
FeaturePlot(object=iHCF, features = "H1_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```
#H2
```{r}
DefaultAssay(iHCF) <- "SCT"
expdata <- GetAssayData(iHCF)
Pop1 <- c("WIF1","DKK3","SFRP2","FRZB")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
iHCF@meta.data$H2_z<-z_scores[1,]
FeaturePlot(object=iHCF, features = "H2_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Injury Response
```{r}
DefaultAssay(iHCF) <- "SCT"
expdata <- GetAssayData(iHCF)
Pop1 <- c("MT1-2","CCL2","CXCL1")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
iHCF@meta.data$InjuryResponse_z<-z_scores[1,]
FeaturePlot(object=iHCF, features = "InjuryResponse_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Myofibroblast
```{r}
DefaultAssay(iHCF) <- "SCT"
expdata <- GetAssayData(iHCF)
Pop1 <- c("ACTA2","CTHRC1","POSTN","FAP")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
iHCF@meta.data$Myofibroblast_z<-z_scores[1,]
FeaturePlot(object=iHCF, features = "Myofibroblast_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Matrifibrocyte
```{r}
DefaultAssay(iHCF) <- "SCT"
expdata <- GetAssayData(iHCF)
Pop1 <- c("COMP","CHAD","CLIP2")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
iHCF@meta.data$Matrifibrocyte_z<-z_scores[1,]
FeaturePlot(object=iHCF, features = "Matrifibrocyte_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

# Late Resolution
```{r}
DefaultAssay(iHCF) <- "SCT"
expdata <- GetAssayData(iHCF)
Pop1 <- c("COL8A1","MEOX1")
pops<-list(Pop1)
#Z-Scores
z_scores<-NULL

for (i in 1:length(pops)) {
genes <- pops[[i]]
zz <- which(tolower(rownames(expdata)) %in% tolower(genes))
av <- numeric(ncol(expdata))

geneExp <- as.matrix(expdata[zz, ])
geneExp <- t(scale(t(geneExp)))
geneExp[is.nan(geneExp)] <- 0
z_scores <- rbind(z_scores,(av + colSums(geneExp) / length(zz)))
}
iHCF@meta.data$LateResolution_z<-z_scores[1,]
FeaturePlot(object=iHCF, features = "LateResolution_z",pt.size=.5, reduction = 'umap') + scale_color_gradientn(colors=c("blue","turquoise2","yellow","red","red4"), oob=scales::squish, limits=c(0,1))
```

```{r}
iHCF$group <- factor(iHCF$group, levels = c("iHCF_72h_TGFb", "iHCF_72h_IL1b", "iHCF_72h_Control"))
```

```{r}
VlnPlot(iHCF, features = c("H1_z","H2_z","InjuryResponse_z","Myofibroblast_z","Matrifibrocyte_z"), group.by = "group", pt.size = 0, stack = TRUE, cols = pal_npg(palette = c("nrc"), alpha = 1)(5)) + NoLegend()
```

```{r}
NHCF <- readRDS("./NHCF/mapping/NHCF_mapped.rds")
NHDF <- readRDS("./NHDF/mapping/NHDF_mapped.rds")
iHCF <- readRDS("./iHCF/mapping/iHCF_mapped.rds")
```

```{r}
VlnPlot(NHCF, features = c("predicted.celltype.l1.score"), group.by = "predicted.celltype.l1", pt.size = 0, cols = c("#D51F26","#272E6A","#208A42","#89288F","#C06CAB","#D8A767","#89C75F")) + NoLegend()
```

```{r}
VlnPlot(NHDF, features = c("predicted.celltype.l1.score"), group.by = "predicted.celltype.l1", pt.size = 0, cols = c("#D51F26","#272E6A","#208A42","#89288F","#FEE500","#8A9FD1","#C06CAB","#D8A767","#90D5E4","#F37B7D")) + NoLegend()
```

```{r}
VlnPlot(iHCF, features = c("predicted.celltype.l1.score"), group.by = "predicted.celltype.l1", pt.size = 0, cols = c("#D51F26","#272E6A","#208A42","#89288F","#C06CAB","#D8A767","#90D5E4","#89C75F","#9983BD")) + NoLegend()
```

```{r}
mean(NHCF$predicted.celltype.l1.score)
```


```{r}
mean(NHDF$predicted.celltype.l1.score)
```

```{r}
mean(iHCF$predicted.celltype.l1.score)
```


```{r}
sum(NHCF$predicted.celltype.l1.score > 0.8)/length(NHCF$predicted.celltype.l1.score)
```

```{r}
sum(NHDF$predicted.celltype.l1.score > 0.8)/length(NHDF$predicted.celltype.l1.score)
```

```{r}
sum(iHCF$predicted.celltype.l1.score > 0.8)/length(iHCF$predicted.celltype.l1.score)
```

```{r}
write.csv(NHCF@meta.data, file ="./NHCF_meta.csv", quote = FALSE)
write.csv(NHDF@meta.data, file ="./NHDF_meta.csv", quote = FALSE)
write.csv(iHCF@meta.data, file ="./iHCF_meta.csv", quote = FALSE)
```

```{r}
write.csv(NHCF@reductions[["ref.umap"]]@cell.embeddings, file ="./NHCF_ref.umap.csv", quote = FALSE)
write.csv(NHDF@reductions[["ref.umap"]]@cell.embeddings, file ="./NHDF_ref.umap.csv", quote = FALSE)
write.csv(iHCF@reductions[["ref.umap"]]@cell.embeddings, file ="./iHCF_ref.umap.csv", quote = FALSE)
```

```{r}
write.csv(fibroblast@reductions[["umap"]]@cell.embeddings, file ="./human_fibroblast.umap.csv", quote = FALSE)
```

###### Revisions

# Fibroblast_sampled
```{r}
DefaultAssay(global) <- "SCT"
DefaultAssay(fibroblast) <- "SCT"
```

```{r}
Idents(global) <- "annotation.0.1"
fibroblast_global <- subset(global, idents = "Fibroblast")
```

```{r}
fibroblast_sampled <- subset(x = fibroblast_global, downsample = 5000)
```

```{r}
anchors_fibroblast_sampled <- FindTransferAnchors(
  reference = fibroblast,
  query = fibroblast_sampled,
  normalization.method = "SCT",
  reference.reduction = "spca",
  dims = 1:50
)
```

```{r}
fibroblast_sampled <- MapQuery(
  anchorset = anchors_fibroblast_sampled,
  query = fibroblast_sampled,
  reference = fibroblast,
  refdata = list(
    celltype.l1 = "functional.cluster",
    predicted_CITE = "CITE"
  ),
  reference.reduction = "spca", 
  reduction.model = "umap"
)
```

```{r}
write.csv(fibroblast_sampled@meta.data, file ="./revisions_pos_neg_control/fibroblast_sampled_meta.csv", quote = FALSE)
```

```{r}
mean(fibroblast_sampled$predicted.celltype.l1.score)
```

```{r}
RidgePlot(fibroblast_sampled, features = "predicted.celltype.l1.score")
```


```{r}
global <- readRDS("/Users/jamrute/Library/CloudStorage/Box-Box/Macbook_Files/Grad_School/Primary_Projects/Amgen_Heart/Amgen_ICM_CITEseq/Human_CITEseq/global_object/final_global_annotated.rds")
```

# Endothelium
```{r}
Idents(global) <- "annotation.0.1"
endothelium <- subset(global, idents = "Endothelium")
```

```{r}
endothelium_sampled <- subset(x = endothelium, downsample = 5000)
```

```{r}
anchors_endothelium_sampled <- FindTransferAnchors(
  reference = fibroblast,
  query = endothelium_sampled,
  normalization.method = "SCT",
  reference.reduction = "spca",
  dims = 1:50
)
```

```{r}
endothelium_sampled <- MapQuery(
  anchorset = anchors_endothelium_sampled,
  query = endothelium_sampled,
  reference = fibroblast,
  refdata = list(
    celltype.l1 = "functional.cluster",
    predicted_CITE = "CITE"
  ),
  reference.reduction = "spca", 
  reduction.model = "umap"
)
```

```{r}
write.csv(endothelium_sampled@meta.data, file ="./revisions_pos_neg_control/endothelium_sampled_meta.csv", quote = FALSE)
```

```{r}
mean(endothelium_sampled$predicted.celltype.l1.score)
```

```{r}
RidgePlot(endothelium_sampled, features = "predicted.celltype.l1.score")
```

```{r}
saveRDS(fibroblast_sampled, "./revisions_pos_neg_control/fibroblast_sampled.rds")
saveRDS(endothelium_sampled, "./revisions_pos_neg_control/endothelium_sampled.rds")
```


# Myeloid
```{r}
Idents(global) <- "annotation.0.1"
myeloid <- subset(global, idents = "Myeloid")
```

```{r}
myeloid_sampled <- subset(x = myeloid, downsample = 5000)
```

```{r}
anchors_myeloid_sampled <- FindTransferAnchors(
  reference = fibroblast,
  query = myeloid_sampled,
  normalization.method = "SCT",
  reference.reduction = "spca",
  dims = 1:50
)
```

```{r}
myeloid_sampled <- MapQuery(
  anchorset = anchors_myeloid_sampled,
  query = myeloid_sampled,
  reference = fibroblast,
  refdata = list(
    celltype.l1 = "functional.cluster",
    predicted_CITE = "CITE"
  ),
  reference.reduction = "spca", 
  reduction.model = "umap"
)
```

```{r}
write.csv(myeloid_sampled@meta.data, file ="./revisions_pos_neg_control/myeloid_sampled_meta.csv", quote = FALSE)
```

```{r}
mean(myeloid_sampled$predicted.celltype.l1.score)
```

```{r}
RidgePlot(myeloid_sampled, features = "predicted.celltype.l1.score")
```

```{r}
saveRDS(endothelium_sampled, "./revisions_pos_neg_control/endothelium_sampled.rds")
```


